---
title: "R Notebook"
output: html_notebook
---

Final Report Presentation
Presented by: Lydia González
April 16, 2021

## Performance Comparison of Tree-Based Classification Algorithms to Predict Wheel-Spinning on Students using the ASSISTments Dataset

## ASSISTments and Skill Builder data set

## What is wheel-spinning?
# 0 - learned skill
# 1 - did not learn skill

# Importing libraries
```{r}
library(rpart)
library(rpart.plot)
library(ipred)
library(caret)
library(ModelMetrics)
library(dplyr)
```

# Importing cleaned CSV file
URL for file: https://github.com/dharmangbhavsar/Predicting-Wheel-Spinning/blob/master/Data/cleaned.csv
```{r}

skillBuilderData <- read.csv("~/Downloads/uofa/wheel-spinning/cleaned.csv")
colnames(skillBuilderData)[19] = "wheelSpinning"
skillBuilderData$wheelSpinning <- factor(ifelse(skillBuilderData$wheelSpinning == "1", 1, 0))
```

# Inspect data set
```{r}
names(skillBuilderData)
dim(skillBuilderData)
str(skillBuilderData)
```

# Split data set into training and testing
```{r}
n <- nrow(skillBuilderData)
n_train <- round(0.8 * n)
set.seed(123)
train_indices <- sample(1:n, n_train)

# Subsetting training data frame
skillBuilderTrain <- skillBuilderData[train_indices, ]

# Subsetting testing data frame
skillBuilderTest <- skillBuilderData[-train_indices, ]
```

# Decision Tree
```{r}
treeModel <- rpart(wheelSpinning ~ answer_type + attempt_count + base_sequence_id + bottom_hint + correct + first_action + hint_count + hint_total + ms_first_response + opportunity + opportunity_original + order_id + overlap_time + skill_id + tutor_mode + user_id, skillBuilderTrain, method = "class")
rpart.plot(treeModel, yesno = 2)

## Generate predictions and evaluate model
test_prediction = predict(treeModel, newdata = skillBuilderTest, type = "class")
caret::confusionMatrix(test_prediction, as.factor(skillBuilderTest$wheelSpinning))

## Generate predicted values on test set and evaluate on AUC
probPrediction <- predict(object = treeModel, newdata = skillBuilderTest, type = "prob")
# Compute AUC
auc(actual = ifelse(skillBuilderTest$wheelSpinning == "1", 1, 0), predicted = probPrediction[, "1"])
```
```{r}
probPrediction
```

```{r}
## Generate predictions and evaluate model
test_prediction1 = predict(treeModel, newdata = skillBuilderTrain, type = "class")
caret::confusionMatrix(test_prediction1, as.factor(skillBuilderTrain$wheelSpinning))


## Generate predicted values on test set and evaluate on AUC
probPrediction1 <- predict(object = treeModel, newdata = skillBuilderTrain, type = "prob")
# Compute AUC
auc(actual = ifelse(skillBuilderTrain$wheelSpinning == "1", 1, 0), predicted = probPrediction1[, "1"])
```


```{r}
df <- data.frame(imp = treeModel$variable.importance)
df2 <- df %>%
  tibble::rownames_to_column() %>% 
  dplyr::rename("variable" = rowname) %>% 
  dplyr::arrange(imp) %>%
  dplyr::mutate(variable = forcats::fct_inorder(variable))
ggplot2::ggplot(df2) +
  geom_col(aes(x = variable, y = imp),
           col = "black", show.legend = F) +
  coord_flip()  +
  scale_fill_gradient2(low = "green", high = "yellow")
```
```{r}
round(treeModel$variable.importance, 2)
```

# Bagged Tree
```{r}
# Change NA values to 0
skillBuilderTrain[is.na(skillBuilderTrain)] <- 0
skillBuilderTest[is.na(skillBuilderTest)] <- 0

baggedModel <- bagging(wheelSpinning ~ answer_type + attempt_count + base_sequence_id + bottom_hint + correct + first_action + hint_count + hint_total + ms_first_response + opportunity + opportunity_original + order_id + overlap_time + skill_id + tutor_mode + user_id, skillBuilderTrain, coob = TRUE)

# Generate predictions and evaluate model
bagged_prediction <- predict(baggedModel, skillBuilderTest, type = "class")
caret::confusionMatrix(bagged_prediction, skillBuilderTest$wheelSpinning)

## Generate predicted values on test set and evaluate on AUC
baggedProbPrediction <- predict(baggedModel, skillBuilderTest, type = "prob")
# Compute AUC
auc(actual = ifelse(skillBuilderTest$wheelSpinning == "1", 1, 0), predicted = baggedProbPrediction[, "1"])
```
```{r}
# Generate predictions and evaluate model
bagged_prediction1 <- predict(baggedModel, skillBuilderTrain, type = "class")
caret::confusionMatrix(bagged_prediction1, skillBuilderTrain$wheelSpinning)

## Generate predicted values on test set and evaluate on AUC
baggedProbPrediction1 <- predict(baggedModel, skillBuilderTrain, type = "prob")
# Compute AUC
auc(actual = ifelse(skillBuilderTrain$wheelSpinning == "1", 1, 0), predicted = baggedProbPrediction1[, "1"])
```
```


```{r}
pred.imp <- varImp(baggedModel)
pred.imp
```


```{r}
barplot(pred.imp$Overall,
        names.arg = row.names(pred.imp))
```
# Discussion
```{r}
df <- data.frame(imp = pred.imp$Overall)
df2 <- df %>%
  tibble::rownames_to_column() %>% 
  dplyr::rename("variable" = rowname) %>% 
  dplyr::arrange(imp) %>%
  dplyr::mutate(variable = forcats::fct_inorder(variable))
ggplot2::ggplot(df2) +
  geom_col(aes(x = variable, y = imp),
           col = "black", show.legend = F) +
  coord_flip()  +
  scale_fill_gradient2(low = "green", high = "yellow")
```
